---
format:
  gfm:
    default-image-extension: ""
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vsn

**V**ariance **s**tabilization and calibration for microarray data.

`vsn` implements a method for normalising single- and multiple-color microarray
intensities (and, in principle, data from other technologies with a similar
format). The model incorporates data calibration step (a.k.a. normalization), a
model for the dependence of the variance on the mean intensity and a variance
stabilizing data transformation. Differences between transformed intensities
are analogous to "normalized log-ratios". However, in contrast to the latter, 
their variance is independent of the mean, and they are usually more sensitive
and specific in detecting differential transcription.

## Installation

`vsn` is part of Bioconductor:

```{r}
#| eval: false
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("vsn")
```

## Usage

```{r}
#| message: false
library("vsn")
```

The simplest way to normalize a dataset is:

```{r}
data("kidney")          # ExpressionSet of unnormalised data
xnorm <- justvsn(kidney)
xnorm
```

For more control, fit the model and apply it in two steps:

```{r}
fit <- vsn2(kidney)
ynorm <- predict(fit, kidney)
ynorm
```

Both are equivalent. The two-step form is useful when fitting on a subset
(e.g. spike-in or control features) and then applying the fit to the full
data, or when you want to inspect the `fit` object. `justvsn()` and `vsn2()`
are also available for `AffyBatch` and `RGList` objects.

## References

```{r}
citation(package = "vsn")
```

## Feedback

This is an approved de.NBI service. Please help us improve by taking our [short user survey](https://www.surveymonkey.de/r/denbi-service?sc=hd-hub&tool=vsn).
